### 2020-10-15 priority-cp.tex

\subsubsection*{Example}\label{sec:pcp-example}
We introduce PCP by revising Milner's cyclic scheduler~\cite{milner89,dardhagay18}. Our PCP scheduler has exactly the same behaviour as the version presented in~\cref{sec:pgv-example}. We use $\pcp{\tm{\prod_{i}P_i}}$ to denote the parallel composition of the processes $\pcp{\tm{P_i}}$, and $\pcp{\plug{P}{Q}}$ to denote the plugging of $\pcp{\tm{Q}}$ in the one-hole process context $\pcp{\tm{P}}$:
\[
  \usingnamespace{pcp}
  \begin{array}{lrlr}
    \tm{\sched}
    & \defeq & \tm{\res{a_1}{a'_1}{\dots{\res{a_n}{a'_n}{}}}}
               \tm{\res{b_1}{b'_1}{\dots{\res{b_n}{b'_n}{}}}}
               \tm{\res{c_1}{c'_1}{\dots{\res{c_n}{c'_n}{}}}}
               \tm{\res{d}{d'}{}}
    \\ &     & \tm{(
               \ppar{\proc1}{\agent1}
               \parallel
               \prod_{2\leq{i}\leq{n}}(\ppar{\proc{i}}{\wait{c'_{i-1}}{\agent{i}}})
               )}
    \\
    \tm{\agent1}
    & \defeq & \tm{\close{a_i}{\wait{b_i}{\close{c_i}{\wait{c'_n}{\close{d}{\halt}}}}}}
    \\
    \tm{\agent{i}}
    & \defeq & \tm{\close{a_i}{\wait{b_i}{\close{c_i}{\halt}}}}
    \\
    \tm{\proc{i}}
    & \defeq & \tm{\wait{a'_i}{\plug{P_i}{\close{b'_i}{Q_i}}}}
  \end{array}
\]

>>>>

\subsubsection*{Example}\label{sec:pcp-example}
We introduce PCP by revising Milner's cyclic scheduler~\cite{milner89,dardhagay18}. Our PCP scheduler has exactly the same behaviour as the version presented in~\cref{sec:pgv-example}. We use $\pcp{\tm{\prod_{i}P_i}}$ to denote the parallel composition of the processes $\pcp{\tm{P_i}}$, and $\pcp{\plug{P}{Q}}$ to denote the plugging of $\pcp{\tm{Q}}$ in the one-hole process-context $\pcp{\tm{P}}$:
\[
  \usingnamespace{pcp}
  \begin{array}{lrlr}
    \tm{\sched}
    & \defeq & \tm{\res{a_1}{a'_1}{\dots{\res{a_n}{a'_n}{}}}}
               \tm{\res{b_1}{b'_1}{\dots{\res{b_n}{b'_n}{}}}}
               \tm{\res{c_1}{c'_1}{\dots{\res{c_n}{c'_n}{}}}}
               \tm{\res{d}{d'}{}}
    \\ &     & \tm{(
               \ppar{\proc1}{\agent1}
               \parallel
               \prod_{2\leq{i}\leq{n}}(\ppar{\proc{i}}{\wait{c'_{i-1}}{\agent{i}}})
               )}
    \\
    \tm{\agent1}
    & \defeq & \tm{\close{a_i}{\wait{b_i}{\close{c_i}{\wait{c'_n}{\close{d}{\halt}}}}}}
    \\
    \tm{\agent{i}}
    & \defeq & \tm{\close{a_i}{\wait{b_i}{\close{c_i}{\halt}}}}
    \\
    \tm{\proc{i}}
    & \defeq & \tm{\wait{a'_i}{\plug{P_i}{\close{b'_i}{Q_i}}}}
  \end{array}
\]

### 2020-10-15 priority-gv.tex

\begin{example}[Milner's Cyclic Scheduler in PGV]\label{sec:pgv-example}
Suppose a set of processes $\tm{\proc{i}}$, $1\leq{i}\leq{n}$, is scheduled to perform some tasks in cyclic order, starting with $\tm{\proc0}$, ending with $\tm{\proc{n}}$, and notifying $\tm{\proc0}$ when all processes have finished.
% The dotted lines represent the optional transfer of data between processes. With three processes, the scheduler looks as follows:\simon{Sorry, I found this diagram super confusing\dots Is there any way you could make it look more cyclic? Also, have the convention of a single annotation per channel, with the convention that the greater agent / proc gets the primed end of the previous channel?}
% \input{fig/pgv-scheduler.tex}
We implement the scheduler as follows, using $\tm{\prod_{i}\conf{C}_i}$ to denote the parallel composition of the processes $\tm{\conf{C}_i}$, and $\tm{\plug{M}{N}}$ to denote the plugging of $\tm{N}$ in the one-hole term context $\tm{M}$:
\[
  \begin{array}{lrlr}
    \tm{\sched}
    & \defeq & \tm{\res{a_1}{a'_1}{\dots{\res{a_n}{a'_n}{}}}}
               \tm{\res{b_1}{b'_1}{\dots{\res{b_n}{b'_n}{}}}}
               \tm{\res{c_1}{c'_1}{\dots{\res{c_n}{c'_n}{}}}}
               \tm{\res{d}{d'}{}}
    \\ &     & \tm{\begin{array}{lll}
                     (
                     & \phi\;\proc1
                     \parallel
                     \child\;\andthen{\agent1}{\andthen{\wait\;{c'_3}}{\close\;{d}}}
                     \\
                     \parallel
                     &
                     \prod_{2\leq{i}\leq{n}}
                     (\child\;\proc{i} \parallel \child\;\andthen{\wait\;{c'_{i-1}}}{\agent{i}})
                     & )
                   \end{array}}
    \\
    \tm{\agent{i}}
    & \defeq & \tm{\andthen{\close\;{a_i}}{\andthen{\wait\;{b_i}}{\close\;{c_i}}}}
    \\
    \tm{\proc{i}}
    & \defeq & \tm{\andthen{\wait\;{a'_i}}{\plug{M_i}{\andthen{\close\;{b'_i}}{N_i}}}}
  \end{array}
\]
Our scheduler $\tm{\sched}$ consists of set of agents $\tm{\agent{i}}$, $1\leq{i}\leq{n}$, each representing their respective process. Each process $\tm{\proc{i}}$ waits for the signal to start their task on $\tm{a'_i}$, and signals completion on $\tm{b'_i}$. Each agent signals their process to start on $\tm{a_i}$, waits for their process to finish on $\tm{b_i}$, and then signals for the next agent to continue on $\tm{c_i}$. The agent $\tm{\agent0}$ initiates, then waits for every other process to finish, and signals $\tm{\proc0}$ on $\tm{d}$. Every other agent $\tm{\agent{i}}$, $2\leq{i}\leq{n}$ waits on $\tm{c'_{i-1}}$ for the signal to start.
Each of the channels in the scheduler is of a terminated type, and is merely used to synchronise.
For simplicity, we let $\tm{\sched}$ be a configuration. However, we could equivalently have represented it as a function which takes three processes as arguments, and creates this configuration.
\end{example}

\subsubsection*{Types}
Types in PGV are divided into session types ($\ty{S}$) and types ($\ty{T}$, $\ty{U}$).
%
Session types are defined by the following grammar:
\[
\begin{array}{lcl}
  \ty{S}
  & \Coloneqq & \ty{\tysend[\cs{o}]{T}{S}}
    \sep        \ty{\tyrecv[\cs{o}]{T}{S}}
    \sep        \ty{\tyends[\cs{o}]}
    \sep        \ty{\tyendr[\cs{o}]}
\end{array}
\]
The session types $\ty{\tysend[\cs{o}]{T}{S}}$ and $\ty{\tyrecv[\cs{o}]{T}{S}}$ describe the endpoints of a channel over which we send or receive a value of type $\ty{T}$, and then proceed as $\ty{S}$. The types $\ty{\tyends[\cs{o}]}$ and $\ty{\tyendr[\cs{o}]}$ describe endpoints of a channel whose communication has finished, and over which we must synchronise before closing the channel. Each connective in a session type is annotated with a \emph{priority} $\cs{o}\in\mathbb{N}$.

Types are defined by the following grammar:
\[
\begin{array}{lcl}
  \ty{T}, \ty{U}
  & \Coloneqq & \ty{\typrod{T}{U}}
    \sep        \ty{\tyunit}
    \sep        \ty{\tysum{T}{U}}
    \sep        \ty{\tyvoid}
    \sep        \ty{\tylolli[\cs{p},\cs{q}]{T}{U}}
    \sep        \ty{S}
\end{array}
\]
The types $\ty{\typrod{T}{U}}$, $\ty{\tyunit}$, $\ty{\tysum{T}{U}}$, and $\ty{\tyvoid}$ are the standard linear $\lambda$-calculus product type, unit type, sum type, and empty type.
The type $\ty{\tylolli[\cs{p},\cs{q}]{T}{U}}$ is the standard linear function type, annotated with \emph{priority bounds} $\cs{p},\cs{q}\in\mathbb{N}\cup\{\cs{\pbot},\cs{\ptop}\}$.
Every session type is also a type.

>>>>

\begin{example}[Milner's Cyclic Scheduler in PGV]\label{sec:pgv-example}
Suppose a set of processes $\tm{\proc{i}}$, $1\leq{i}\leq{n}$, is scheduled to perform some tasks in cyclic order, starting with $\tm{\proc0}$, ending with $\tm{\proc{n}}$, and notifying $\tm{\proc0}$ when all processes have finished.
% The dotted lines represent the optional transfer of data between processes. With three processes, the scheduler looks as follows:\simon{Sorry, I found this diagram super confusing\dots Is there any way you could make it look more cyclic? Also, have the convention of a single annotation per channel, with the convention that the greater agent / proc gets the primed end of the previous channel?}
% \input{fig/pgv-scheduler.tex}
We implement the scheduler as follows, using $\tm{\prod_{i}\conf{C}_i}$ to denote the parallel composition of the processes $\tm{\conf{C}_i}$, and $\tm{\plug{M}{N}}$ to denote the plugging of $\tm{N}$ in the one-hole term-context $\tm{M}$:
\[
  \begin{array}{lrlr}
    \tm{\sched}
    & \defeq & \tm{\res{a_1}{a'_1}{\dots{\res{a_n}{a'_n}{}}}}
               \tm{\res{b_1}{b'_1}{\dots{\res{b_n}{b'_n}{}}}}
               \tm{\res{c_1}{c'_1}{\dots{\res{c_n}{c'_n}{}}}}
               \tm{\res{d}{d'}{}}
    \\ &     & \tm{\begin{array}{lll}
                     (
                     & \phi\;\proc1
                     \parallel
                     \child\;\andthen{\agent1}{\andthen{\wait\;{c'_3}}{\close\;{d}}}
                     \\
                     \parallel
                     &
                     \prod_{2\leq{i}\leq{n}}
                     (\child\;\proc{i} \parallel \child\;\andthen{\wait\;{c'_{i-1}}}{\agent{i}})
                     & )
                   \end{array}}
    \\
    \tm{\agent{i}}
    & \defeq & \tm{\andthen{\close\;{a_i}}{\andthen{\wait\;{b_i}}{\close\;{c_i}}}}
    \\
    \tm{\proc{i}}
    & \defeq & \tm{\andthen{\wait\;{a'_i}}{\plug{M_i}{\andthen{\close\;{b'_i}}{N_i}}}}
  \end{array}
\]
Our scheduler $\tm{\sched}$ consists of set of agents $\tm{\agent{i}}$, $1\leq{i}\leq{n}$, each representing their respective process. Each process $\tm{\proc{i}}$ waits for the signal to start their task on $\tm{a'_i}$, and signals completion on $\tm{b'_i}$. Each agent signals their process to start on $\tm{a_i}$, waits for their process to finish on $\tm{b_i}$, and then signals for the next agent to continue on $\tm{c_i}$. The agent $\tm{\agent0}$ initiates, then waits for every other process to finish, and signals $\tm{\proc0}$ on $\tm{d}$. Every other agent $\tm{\agent{i}}$, $2\leq{i}\leq{n}$ waits on $\tm{c'_{i-1}}$ for the signal to start.
Each of the channels in the scheduler is of a terminated type, and is merely used to synchronise.
For simplicity, we let $\tm{\sched}$ be a configuration. However, we could equivalently have represented it as a function which takes three processes as arguments, and creates this configuration.
\end{example}

\subsubsection*{Session types}
Session ($\ty{S}$) types are defined by the following grammar:
\[
\begin{array}{lcl}
  \ty{S}
  & \Coloneqq & \ty{\tysend[\cs{o}]{T}{S}}
    \sep        \ty{\tyrecv[\cs{o}]{T}{S}}
    \sep        \ty{\tyends[\cs{o}]}
    \sep        \ty{\tyendr[\cs{o}]}
\end{array}
\]
The session types $\ty{\tysend[\cs{o}]{T}{S}}$ and $\ty{\tyrecv[\cs{o}]{T}{S}}$ describe the endpoints of a channel over which we send or receive a value of type $\ty{T}$, and then proceed as $\ty{S}$. The types $\ty{\tyends[\cs{o}]}$ and $\ty{\tyendr[\cs{o}]}$ describe endpoints of a channel whose communication has finished, and over which we must synchronise before closing the channel. Each connective in a session type is annotated with a \emph{priority} $\cs{o}\in\mathbb{N}$.

\subsubsection*{Types}
Types ($\ty{T}$, $\ty{U}$) are defined by the following grammar:
\[
\begin{array}{lcl}
  \ty{T}, \ty{U}
  & \Coloneqq & \ty{\typrod{T}{U}}
    \sep        \ty{\tyunit}
    \sep        \ty{\tysum{T}{U}}
    \sep        \ty{\tyvoid}
    \sep        \ty{\tylolli[\cs{p},\cs{q}]{T}{U}}
    \sep        \ty{S}
\end{array}
\]
The types $\ty{\typrod{T}{U}}$, $\ty{\tyunit}$, $\ty{\tysum{T}{U}}$, and $\ty{\tyvoid}$ are the standard linear $\lambda$-calculus product type, unit type, sum type, and empty type.
The type $\ty{\tylolli[\cs{p},\cs{q}]{T}{U}}$ is the standard linear function type, annotated with \emph{priority bounds} $\cs{p},\cs{q}\in\mathbb{N}\cup\{\cs{\pbot},\cs{\ptop}\}$.
Every session type is also a type.
